### 2020-08-18 text/article_vedrovice.Rmd
improve figures and tables

<!-- sample (burials, artefacts) -->
Originally, the burial ground consisted of at least 108 burials.
Some of the burials were disturbed by later settlement and burial pits in the
area of the cemetery, amateur excavations during the 19th century or later
agricultural activities.
These burials are excluded from the analysis due to missing and/or
incomplete data and only a sample of 78 well-preserved burials is analyzed here.
Most of the data was obtained from the original publication [@podborsky2002],
anthropological determination of sex and age were corrected based on
more recent analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.
Further details on use and function of the artefacts were gathered from
@masclamasclanslatorre2020.

@@

To explore whether neighboring individuals are more often of the same sex we use
two methods to establish neighborhood.
The first method is based on constructing a Gabriel graph [@gabriel1969]
with burial locations as nodes (Fig. \@ref(fig:gabriel)).
Gabriel graph is a subgraph of a Delaunay triangulation

For each node, mean sex of the neighbors is determined and recorded.
To check how common the observed value is, a permutation test
where sex of burials is randomly assigned while preserving the overall
proportion of given categories and the neighborhood struture determined by the
Gabriel graph is performed.
The observed values of neighbors sex categories are compared against distributions
derived from the simulated data.

```{r gabriel, fig.cap = "Gabriel graph"}
knitr::include_graphics(here("plots", "plan_gabriel.pdf"))
```

The second approach of establishing neigborhood is by constructing buffer zones
around each of the burials with 6m radius.
The size of a radius is based on the results of nearest neighbors analysis, where
at the given radius there is strong evidence of a clustering process (Fig. \ref(fig:spat-funs)).
The randomness is again tested using a permutation test similar to the one
described for the neighborhood derived from the Gabriel graph.

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcore]  using functions of many different packages, for the complete list
see the supplementary information.
<!-- readr [@wickham2017a], dplyr [@wickham2016] and
reshape2 [@wickham2007] packages to manipulate the data.
Function daisy [@kaufman1990] from package cluster [@maechler2016] is used to
count distances between observations used in further analyses. -->

@@

```{r label-spat-funs}
label_spat_funs <- paste("Functions *G*, *F*, *K* and *T*.",
"*Dashed line* -- theoretical curve under complete spatial randomness, *solid line* -- empirical curve, *gray area* -- envelope.")
```
```{r spat-funs, fig.cap = label_spat_funs}
knitr::include_graphics(here("plots", "pointprocess_fun.pdf"))
```

```{r label-nb}
label_neigh_sex <- paste("Mean neighbors in a given sex category",
"*Gray area* -- density estimation for the simulated data, *vertical line* -- observed value.",
"*F* -- female, *M* -- male, *n. a.* -- non adult.")
```
```{r nb-buffer, fig.cap = label_neigh_sex}
knitr::include_graphics(here("plots", "nb_buffer_sex.pdf"))
```
```{r nb-gabriel, fig.cap = label_neigh_sex}
knitr::include_graphics(here("plots", "nb_gabriel_sex.pdf"))
```

@@

```{r f-sstat, fig.cap = "Distribution of simulated and observed S statistic"}
knitr::include_graphics(here("plots", "s_statistic.pdf"))
```

@@

>>>>

<!-- sample (burials, artefacts) -->
Originally, the burial ground consisted of at least 108 burials.
Some of the burials were disturbed by later settlement and burial pits in the
area of the cemetery, amateur excavations during the 19th century or later
agricultural activities.
These burials are excluded from the analysis due to missing and/or
incomplete data and only a sample of 78 well-preserved burials is analyzed here.
Most of the data was obtained from the original publication [@podborsky2002],
anthropological determination of sex and age were corrected based on
more recent analysis by @dockalova2008. Carbon, nitrogen and strontium isotope
ratios were collected in @whittle2013 and @richards2008.
Further details on use and function of the artefacts were gathered from
@masclanslatorre2020.

@@

To explore whether neighboring individuals are more often of the same sex we use
three methods to establish neighborhood relations.
The first two methods are based on constructing neighborhood graphs.
We use Delaunay triangulation and its subgraph, a Gabriel graph [@gabriel1969]
with burial locations as nodes (Fig. \@ref(fig:plan-graphs)).
The Delaunay triangulation can be considered a maximum number of neighbors case
scenario while the Gabriel graph the exact opposite.
For each node, mean number of neighbors of a given sex is determined and recorded.
To check how common the observed value is, a permutation test
where sex of burials is randomly assigned while preserving the overall
proportion of given categories and the neighborhood structure determined by the
graphs is performed.
The observed values of neighbors sex categories are compared against distributions
derived from the simulated data.

The third approach of establishing neigborhood is by constructing buffer zones
around each of the burials with 6 meters radius.
The size of a radius is based on the results of nearest neighbors analysis, where
at the given radius there is strong evidence of a clustering process (Fig. \@ref(fig:spat-funs)).
The randomness is again tested using a permutation test similar to the one
described for the neighborhood derived from the graphs.

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcore] using functions of many different packages, for the complete list
see the supplementary information.

<!-- readr [@wickham2017a], dplyr [@wickham2016] and
reshape2 [@wickham2007] packages to manipulate the data.
Function daisy [@kaufman1990] from package cluster [@maechler2016] is used to
count distances between observations used in further analyses. -->

@@

## Neighborhood structure

Three methods of neighborhood relationship construction were examined.
One is distance based where all the neighbors that fall into a given buffer are considered,
the other two are based on neighborhood graphs, the Delaunay triangulation as a
maximum number of neighbors scenario and a Gabriel graph as a minimal scenario.

For relationships (see Fig. \@ref(fig:nb-sex) and Tab. \@ref(tab:nb-tab-sex)) of male-sexed bodies, none of the methods showed any significant
differences between expected and observed number of neighbors of a given sex.
The relationship with female-sexed bodies is rather extreme for all of the methods employed,
but never significant ($p > 0.07$).
In case of bodies unsexed due to early age (non-adults), the neigborhood based
on buffer zones show a signficant realtionship with other non-adults ($p \approx 0.04$).
Finally burials with female-sexed bodies show a significantly higher number of
burials with male-sexed bodies in both the methods based on neighborhood networks
(Delaunay triangulation $p \approx 0.02$ and Gabriel graph $p \approx 0.03$).
Two further relationships showed non-random deviations.
In case of Delaunay trinagulation for burials with bodies with indeterminate sex with
non-adults ($p \approx 0.02$) and in case of Gabriel graph between non-adults and
indeterminate category $p \approx 0.02$.

@@

<!-- ======================================================================= -->
<!-- FIGURES --><!-- FIGURES --><!-- FIGURES --><!-- FIGURES --><!-- FIGURES -->
```{r plan-graphs, fig.cap = "Neighborhood graphs: Delaunay triangulation and Gabriel graph"}
knitr::include_graphics(here("plots/plan_gabriel.pdf"))
```

```{r label-spat-funs}
label_spat_funs <- paste("Functions *G*, *F*, *K* and *T*.",
"*Dashed line* -- theoretical curve under complete spatial randomness, *solid line* -- empirical curve, *gray area* -- envelope.")
```
```{r spat-funs, fig.cap=label_spat_funs}
knitr::include_graphics(here("plots/pointprocess_fun.pdf"))
```

```{r label-nb-sex}
label_nb_sex <- paste("Mean neighbors in a given sex category.",
"*Gray area* -- density estimation for the simulated data, *vertical line* -- observed value.",
"*F* -- female-sexed body, *M* -- male-sexed body, *n. a.* -- non adult, *ind.* -- indetermined sex.")
```
```{r nb-sex, fig.cap=label_nb_sex}
knitr::include_graphics(here("plots/nb_sex.pdf"))
```

```{r f-sstat, fig.cap = "Distribution of simulated and observed S statistic"}
knitr::include_graphics(here("plots/s_statistic.pdf"))
```

@@

<!-- ======================================================================= -->
<!-- TABLES --><!-- TABLES --><!-- TABLES --><!-- TABLES --><!-- TABLES -->
```{r nb-tab-sex}
tab <- read_csv(here("data/temp/nb_sex_p.csv")) %>%
         mutate(relation = str_c(from, to, sep = " - ")) %>%
  select(relation, everything(), -from, -to)

tnames <- c("relation", rep(c("mean", "p"), 3))

knitr::kable(tab, booktabs = TRUE, col.names = tnames, linesep = "", caption = "Results...") %>%
  kableExtra::add_header_above(c(" ", "Buffer" = 2, "Delaunay" = 2, "Gabriel" = 2))
```

### 2020-08-18 text/si_vedrovice.Rmd
improve figures and tables

@@

>>>>

```{r}
library(here)
library(tidyverse)
library(knitr)
library(kableExtra)
```

# Spatial analysis

<!--
If $s_i$ is an event location, then
$G$ function is given by the formula
\@ref(eq:gfun)

\begin{equation}
\label{eq:gfun}
G(d) = \frac{\#[d_{min}(s_i)<d]}{n}
\end{equation}

$F$ function by
$$F(d) = \frac{\#[d_{min}(p_i, S)<d]}{m}$$
where ${p_1 \cdots p_i \cdots p_m}$ is a set of $m$ randomly selected locations.
Ripley's $K$ function
$$K(d) = \frac{\alpha}{n}\frac{1}{n}\sum_{i=1}^n \#[S \in C(s_i, d)]$$
$T$ function
$$ $$
-->

$s_i$ is an event location.

## G function

$$ G(d) = \frac{\sum[d_{min}(s_i)<d]}{n} $$

## F function

$$F(d) = \frac{\#[d_{min}(p_i, S)<d]}{m}$$

where ${p_1 \cdots p_i \cdots p_m}$ is a set of $m$ randomly selected locations.

## Ripley's K function

$$K(d) = \frac{\alpha}{n}\frac{1}{n}\sum_{i=1}^n \#[S \in C(s_i, d)]$$

@@

# Results

```{r, echo = FALSE, message = FALSE, results = 'asis'}
tab <- read_csv(here("data/temp/nb_sex_p.csv")) %>% 
         mutate(relation = str_c(from, to, sep = " - ")) %>% 
  select(relation, everything(), -from, -to)

tnames <- c("relation", rep(c("mean", "p"), 3))

kable(tab, booktabs = TRUE, col.names = tnames) %>% 
  add_header_above(c(" ", "Buffer" = 2, "Delaunay" = 2, "Gabriel" = 2))
```
